---
title: "Tidy data and joins"
author: "Rachel Han"
date: "07/10/2019"
output:
  html_document:
    highlight: tango
    number_sections: yes
    theme: united
    toc: yes
    toc_float: yes
    keep_md: yes
  github_document:
    toc: yes
  pdf_document:
    toc: yes
---
```{r, echo=FALSE, warning=FALSE,include=FALSE}
library(tidyverse)
library(gapminder)
library(ggplot2)
library(knitr)
library(DT)
library(rmarkdown)
library(ggridges)
library(tsibble)
library(scales)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Univariate data reshaping - Option 1

- Make a tibble with one row per year, and columns for life expectancy for two or more countries.


```{r}
(gapminder_wide <- gapminder %>% filter(country %in% c("Canada","Korea, Rep.")) %>%
   pivot_wider(id_cols= year,
              names_from = country,
              values_from = lifeExp)) %>% datatable()
```

-Take advantage of this new data shape to scatterplot life expectancy for one country against that of another.

```{r}
gapminder_wide %>% ggplot(aes(Canada, `Korea, Rep.`)) + geom_point(aes(color=year)) +
  ggtitle("Life expectancy comparison") +
  xlab("life expectancy of Canada") +
  ylab("life expectancy of S. Korea")
```

-Re-lengthen the data.

```{r}
gapminder_wide %>% pivot_longer(cols = -year,
                                names_to ="country",
                                values_to = "lifeExp") %>% datatable()
```

# Multivariate Data Reshaping 

- Make a tibble with one row per year, and columns for life expectancy and GDP per capita (or two other numeric variables) for two or more countries.

```{r}
(gapminder_multi <- gapminder %>%
   filter(country %in% c("Canada","Korea, Rep.")) %>% 
   pivot_wider(id_cols = year,
               names_from = country,
               values_from = c(lifeExp, gdpPercap))) %>% datatable()
```

- Re-lengthen the data.

```{r}
gapminder_multi %>% pivot_longer(cols=-year,
                                 names_to = c(".value","country"),
                                 names_sep = "_") %>% datatable()
```

# Table Joins
Loading data from web...
```{r echo = T, include = FALSE}
guest <- read_csv("https://raw.githubusercontent.com/STAT545-UBC/Classroom/master/data/wedding/attend.csv")
email <- read_csv("https://raw.githubusercontent.com/STAT545-UBC/Classroom/master/data/wedding/emails.csv")
```

Let's view the data:

```{r}
guest %>% datatable()
```

```{r}
email %>% kable()
```


We notice that the multiple people have same emails (one family gave one email). We fix this:

```{r}
(email_fixed <- email %>% 
   separate_rows(guest, sep = ", ")  %>% rename(name=guest)) %>% 
   datatable() 
```

This looks better.


## 3.1 
- For each guest, add a column for email address.
```{r}
(guest_email_join <- left_join(guest,email_fixed,
                              by="name")) %>% datatable()
```

```{r}
guest %>% left_join(email_fixed,by="name") %>% datatable()
```

## 3.2

- Who do we have emails for, yet are not on the guest list?
```{r}
email_fixed %>% anti_join(guest, by = "name") %>% kable()
```

## 3.3

- Make a guestlist that includes everyone we have emails for.

```{r}
guest %>% full_join(email_fixed) %>% datatable()
```

