---
title: 'Homework 1: `mtcars` data exploration'
author: "Rachel Han"
date: "15/09/2019"
output:
  pdf_document:
    toc: yes
  html_document:
    highlight: tango
    number_sections: no
    theme: united
    toc: yes
    toc_float: no
    keep_md: yes
---
##  `mtcars` as tibble


```{r, echo=FALSE, warning=FALSE,include=FALSE}
library(tidyverse)
library(gapminder)
library(tibble)
```

`mtcars` data set is of type data.frame, not a tibble. A tibble lets us see the data without printing the entire data set to the console. An example is shown below:

```{r}
mtcars_tibble <-as_tibble(mtcars)
mtcars_tibble
```

## Dimensions of data
 let's see how many rows and columns are there:
```{r}
rownum <- nrow(mtcars_tibble)
colnum <- ncol(mtcars_tibble)
cat("row:",rownum)
cat("column:",colnum)
```

## Glimpse
We can use the `glimpse` function to see the glimpse of the data. Notice that the data is transposed upon applying this function: the column names are now in rows. 
```{r}
glimpse(mtcars_tibble)
```

## Row names
Notice that by changing the data type to tibble, the row names of `mtcars` are not the strings of car names, but numbers representing the row.

Originally, the row names are:
```{r}
rownames(mtcars)
```
After conversion:
```{r}
rownames(mtcars_tibble)
```

## Summary
We can use the function `summary` to get some basic statistical information about the data.
```{r}
summary(mtcars_tibble)
```

