---
title: '`mtcars` data exploration'
author: "Rachel Han"
date: "15/09/2019"
output:
  beamer_presentation: default
  ioslides_presentation: null
  keep_md: yes
  widescreen: yes
---
## `mtcars` as tibble


```{r, echo=FALSE, warning=FALSE,include=FALSE}
library(tidyverse)
library(gapminder)
library(tibble)
```


`mtcars` data set is of type data.frame, not a tibble. A tibble lets us see the data without printing the entire data set to the console. An example is shown below:

```{r}
mtcars_tibble <-as_tibble(mtcars)
head(mtcars_tibble[,1:5])
```
---
```{r, echo=FALSE, warning=FALSE,include=FALSE}
mtcars_tibble_subset <-as_tibble(mtcars[1:10,1:5])
```

For the purpose of the presentation, we will use a smaller subset of the data.

## Dimensions of data
Let's see how many rows and columns are there:
```{r}
rownum <- nrow(mtcars_tibble)
colnum <- ncol(mtcars_tibble)
cat("row:",rownum)
cat("column:",colnum)
```

## Glimpse {.smaller}
We can use the `glimpse` function to see the glimpse of the data. Notice that the data is transposed upon applying this function: the column names are now in rows. 
```{r}
glimpse(mtcars_tibble[1:4,1:4])
```

## Row names
Notice that by changing the data type to tibble, the row names of `mtcars` are not the strings of car names, but numbers representing the row.

Originally, the row names are:
```{r}
rownames(mtcars[1:3,])
```
---
After conversion:
```{r}
rownames(mtcars_tibble[1:3,])
```


## Summary {.smaller}
We can use the function `summary` to get some basic statistical information about the data. Here, we just select the first five columns.
```{r}
summary(mtcars_tibble[,1:3])
```

