---
title: "Exploring `gapminder` Part 2"
author: "Rachel Han"
date: "30/09/2019"
output:
  github_document:
    toc: yes
  html_document:
    highlight: tango
    number_sections: yes
    theme: united
    toc: yes
    toc_float: yes
    keep_md: yes
  pdf_document:
    toc: yes
---
```{r, echo=FALSE, warning=FALSE,include=FALSE}
library(tidyverse)
library(gapminder)
library(ggplot2)
library(knitr)
library(DT)
library(rmarkdown)
library(ggridges)
library(tsibble)
library(scales)
```

# Exploring Life Expectancies Globally

Let's examine the distribution of life expectancy across all countries.
```{r}
gapminder %>%
  ggplot(aes(lifeExp, color = lifeExp)) +
  geom_histogram(bins = 50, aes(fill=..count..)) +
  scale_fill_distiller(palette = "Greens")
```

Here is a worldwide statistics about life expectancy:


```{r}
gapminder %>% select(lifeExp, year) %>%
  summarise(median_lifeExp = median(lifeExp),
            mean_lifeExp = mean(lifeExp)) %>% kable

```

- The table above shows the median and the mean life expectancy across all the countries in `gapminder`. Being optimistic, let's pick the median life expectancy,`median_lifeExp = 60.7125`, as a global standard. 
- The black vertical line on the histogram denotes the global standard.
```{r}
gapminder %>%
  ggplot(aes(lifeExp, color = lifeExp)) +
  geom_histogram(bins = 50, aes(fill=..count..)) +
  scale_fill_distiller(palette = "Greens") +
  geom_vline(aes(xintercept = median(lifeExp)))
```

- Now, let's find the countries that have life expectancies that are below this global standard.
```{r}
p <- gapminder %>%
  select(country, continent, year, lifeExp) %>%
  filter(lifeExp < median(lifeExp)) %>% 
  group_by(continent, year) %>% 
  summarise(count_countries=n())
(p)
```

Notice Oceania is not here. Why?

```{r}
gapminder %>% filter(continent == "Oceania") %>%
  select(country,year,lifeExp,pop) %>% 
  filter(lifeExp == min(lifeExp)) %>% kable
```


The minimum life expectancy recorded is 69.12 in Australia in 1952, which is higher than the global standard.

## Life Expectancy by Continent

```{r}
p %>% ggplot(aes(x=year, y = count_countries,group = continent, color = continent)) + geom_line() + geom_point() + ylab("Number of countries") + xlab("Year") + ggtitle("Country count per continent with low life expectancy")
```

- Comparing against the global standard (median life expectancy), we see that Africa had the highest count of countries with early death across all times. 
- For Asia, the count decreases rapidly over time.
- Oceania is left out of the graph as no countries are below the median.

# Continent-wise Life Expectancy Trend

```{r}
gapminder %>%
  select(country, continent, year, lifeExp, pop) %>%
  group_by(continent,year) %>%
  summarise(median_lifeExp=median(lifeExp),
            sum_pop =sum(as.numeric(pop))) %>% 
  ggplot(aes(year,median_lifeExp, color = continent))+
  geom_smooth(aes(group=continent)) 
```

We can see that over time, the life expectancy generally goes up. The slowest increase of life expectancy is observed in Europe and Oceania. The fastest rise is in Asia. For Africa, we see that during mid 1980's to 2000, the slope is levelling to a constant.

# Correlations

Could there be a correlation between life expectancy, and population? Below, we plot the the median population versus median life expectancy over time per continent.

We see that in general, as population increases, the life expectancy increases.

```{r}
gapminder %>%
  select(country, continent, year, lifeExp, pop) %>%
  group_by(continent,year) %>%
  summarise(median_lifeExp=median(lifeExp),
            median_pop =median(as.numeric(pop))) %>% 
  ggplot(aes(median_pop,median_lifeExp, color = continent)) +   geom_point(aes(group=continent)) + scale_x_log10(labels = comma_format()) + ggtitle("Median pop vs. Median lifeExp") 

```

We also plot the median GDP per capita versus the median expectancy. We also see that there is an increasing trend.

```{r}
gapminder %>%
  select(country, continent, year, lifeExp, gdpPercap) %>%
  group_by(continent,year) %>%
  summarise(median_lifeExp=median(lifeExp),
            median_gdp =median(gdpPercap)) %>% 
  ggplot(aes(median_gdp,median_lifeExp, color = continent)) +   geom_point(aes(group=continent)) + scale_x_log10(labels = comma_format()) + ggtitle("Median lifeExp vs. Median population") 

```

I will not draw any qualitative interpretations from this data, more statistical analysis should be done in order to do so.


# Weighted mean of life expectancy, by population
```{r}
gapminder %>% group_by(continent,year) %>%
  summarise(vanilla_mean = mean(lifeExp),
            mean_lifeExp_pop = weighted.mean(lifeExp,pop))
```

- `mean_lifeExp_pop` represents the mean of life expectancy weighted by population. This means that the life expectancy data with bigger population gets a higher weight than those with smaller population. 

Here is a plot of the vanilla mean and the weighted mean. The thinner plot denotes the vanilla mean.

```{r}
gapminder %>% group_by(continent, year) %>%
  summarise(vanilla_mean = mean(lifeExp),
            mean_lifeExp_pop = weighted.mean(lifeExp,pop)) %>% ggplot(aes(year,mean_lifeExp_pop,
              color = continent)) +
  geom_line(aes(group= continent)) +
  geom_point(aes(year,mean_lifeExp_pop,
                 group = continent)) +
  geom_line(aes(year,vanilla_mean), size = 0.8) +
  geom_point(aes(year,vanilla_mean,
                 group = continent))
            
```



